---
title: "Introduction to bonsai"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to bonsai}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The goal of bonsai is to provide bindings for additional tree-based model engines for use with the {parsnip} package. 

If you're not familiar with parsnip, you can read more about the package on it's [website](https://parsnip.tidymodels.org).

To get started, load bonsai with:

```{r setup}
library(bonsai)
```

To illustrate how to use the package, we'll fit some models to a dataset containing measurements on 3 different species of penguins. Loading in that data and checking it out:

```{r}
library(modeldata)

data(penguins)

str(penguins)
```

Specifically, we'll make use of measurements on penguin's flippers, and the island that they live on, to predict their species using a decision tree. We'll do so using the engine `"partykit"`—this is one of the packages that bonsai supports that will be used to actually fit model parameter estimates.


```{r}
# set seed for reproducibility
set.seed(1)

# specify and fit model
dec_mod <- 
  decision_tree() %>%
  set_engine(engine = "partykit") %>%
  set_mode(mode = "classification") %>%
  fit(
    formula = species ~ flipper_length_mm + island, 
    data = penguins
  )

dec_mod
```

From this output, we can see that the model generally first looks to `island` to determine species, and then makes use of a mix of flipper length and island to ultimately make a species prediction. There are a total of `r ` terminal nodes in this tree.

```{r}

```
